---
title: "DNS and CDN Consolidation - and the Fundamental Change to the Architecture of the Web"
author:
  - Dan York
  - Thomas Peterson
date: "`r format(Sys.time(), '%d %B %Y')`"
classoption: twocolumn
output: 
  pdf_document:
    keep_tex: true
    toc: no
  html_document:
    toc: yes
---

# Notice

This notebook is intended to be used for collation and initial writeups of the analysis
of our work, and will not look anything like final publication.

---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, results="hide")

library(DBI)
library(ggplot2)

if(!exists("dataDir")) {
  dataDir <- "../../data/"
}

if(!exists("resourcesDir")) {
  resourcesDir <- "../../resources/"
}

mmdb <- system.file("extdata",paste0(resourcesDir, "2019-08-15/", "GeoLite2-ASN.mmdb"), 
                    package = "rgeolocate")


sql_file <- paste0(dataDir, "journal-cyber-policy-2019/", "2018-08-16-snapshot.sqlite")

con <- dbConnect(RSQLite::SQLite(), sql_file)
```

# Abstract

When the Web began, the relationship was between web clients (browsers) and web
servers that existed on a single IP address. Thirty years later, a fundamental
shift has occurred where browsers no longer interact directly with the original
web servers. Instead a significant percentage of websites use a content
distribution network (CDN) or a cloud provider offering CDN/caching capabilities.

Our research began in an attempt to answer a question that arose within the
Domain Name System (DNS) Operations working group in the Internet Engineering
Task Force (IETF) about the actual usage of “CNAME” records in DNS to redirect
the “www” subdomain to another address. In our initial analysis of the Alexa top
1 million domain names, we found 47% of all “www” subdomains were using CNAME
records for some type of redirection. Investigating further, we found a sizeable
percentage of sites hosted by a CDN or cloud provider. Further, these providers
appeared comprised of a small set including Google, Amazon, Akamai, and
Cloudflare.

In our research, we also observed that a significant percentage of the
authoritative name servers for these domains were also in a small of set of
DNS hosting providers, some of which also provided CDN/cloud services.
The emergence of these small sets of providers, and the overlap between the
CDN and DNS providers, caused us to want to dig further to understand the true
level of consolidation within the CDN/cloud/caching infrastructure of the Web
and also within the DNS of the Internet.

Our research and analysis for this paper will focus on these questions:

* How many domains are hosted by known CDNs or cloud providers?
* How much consolidation has occurred within the CDNs and cloud providers?
* Related, how much consolidation has occurred within the DNS hosting providers,
and how does the choice of hosting provider relate to DNS hosting?

# Keywords

DNS; content delivery network; cloud computing

# Introduction

# Materials and Methods

In order to determine both the website hosting provider and the DNS hosting
provider for a given domain, we performed a series of DNS queries - the A, AAAA,
and NS records for both the provided host name, with additional A and AAAA record
queries performed against the NS records. In addition, we performed queries
against the A and AAAA records for the "www" subdomain on each hostname. All
successful responses for A and AAAA records had their IP address's respective
Autonomous System Number looked up against a Maxmind GeoLite2 databased which
was downloaded at the same time as the list of hostnames.

All code, data, analysis and this report are available on GitHub.

# Results

## Identification of DNS providers
We have identified two specific means to identify for a given host name the DNS 
hosting provider, firstly performing matching against [known DNS providers](https://en.wikipedia.org/wiki/List_of_managed_DNS_providers), in
addition to including other known DNS registrars acting as hosting providers, as
well as the hosting providers or CDNs themselves. This process will exclude
scenarios where the host name's delegated nameservers are not assigned to a known
host name.
```{r}
nameServerList <- list(

  # DNS Providers
  c("DNS Provider", "Dyn", "dynect."),
  c("DNS Provider", "NS1", "nsone.net."),
  c("DNS Provider", "Verisign", "verisigndns"),
  c("DNS Provider", "UltraDNS", "ultradns."),
  c("DNS Provider", "DNSimple", "dnsimple.com."),
  c("DNS Provider", "ClouDNS", "cloudns.net."),
  c("DNS Provider", "CDNetworks", "cdnetdns."),
  c("DNS Provider", "CloudfloorDNS", "cloudfloordns."),
  c("DNS Provider", "DNS Made Easy", "dnsmadeeasy.com"),
  c("DNS Provider", "EasyDNS", "easydns."),
  c("DNS Provider", "No-IP", "no-ip.com."),
  c("DNS Provider", "Constellix", "constellix."),

  # Registrars
  c("Registrar", "Namecheap", "registrar-servers.com."),
  c("Registrar", "Gandi", "gandi.net."),
  c("Registrar", "Godaddy", "domaincontrol.com."),
  c("Registrar", "123-reg", "123-reg.co.uk."),
  c("Registrar", "eNom", "name-services.com."),

  # Hosting Providers
  c("Hosting Provider", "AWS Route53", ".awsdns-"),
  c("Hosting Provider", "Azure", "azure-dns"),
  c("Hosting Provider", "Google", "googledomains.com"),
  c("Hosting Provider", "Alibaba", "alibabadns.com."),
  c("Hosting Provider", "Linode", "linode.com."),
  c("Hosting Provider", "DigitalOcean", "digitalocean.com."),
  c("Hosting Provider", "Godaddy", "secureserver.net."),
  c("Hosting Provider", "OVH", "ovh.net."),

  # Content Delivery Networks
  c("Content Delivery Network", "CloudFlare", "cloudflare.com."),
  c("Content Delivery Network", "Akamai", "akam.net"),

  # Content Providers
  c("Content Provider", "Wordpress", "wordpress.com"),
  c("Content Provider", "Google", "google.com."),
  c("Content Provider", "Squarespace", "squarespacedns.com.")
)
```

Another method to identifying DNS hosting providers is to use the autonomous
system number assigned to any IP addresses that are. This may fail in scenarios
where the DNS provider does not have an autonomous system number of their own,
or utilise other networks to host their authoritative servers. For example, 
`qq.com` ranked #14, uses 10 different networks. Despite this, there can be a
correlation between many of the large, notable hosting providers and CDNs and
the autonomous system allocations they use.

```{sql, connection=con, output.var="ns_asn_count"}
-- For each domain, count the total number of distinct ASNs used by the
-- authoritative name servers.
SELECT COUNT(DISTINCT asn) AS ns_asn_total FROM `record`
WHERE record_type IN("A", "AAAA")
AND response_code = 0
AND is_www = 0
AND parent IS NOT NULL
GROUP BY domain
```

```{r}
ggplot(ns_asn_count, aes(ns_asn_total)) +
  geom_freqpoly(binwidth = 1) +
  xlim(1, 13)
```

# Discussion

# Acknowledgments

# Declaration of interest statement

# References